# Results
```{r}
library(ggplot2)
library(tidyr)
library(dplyr)
library(GGally)
```


```{r}

nyc_weekly_earnings <- read.csv("nyc_weekly_earnings.csv")

for (col in c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov"
              , "Dec", "Annual")) {
  nyc_weekly_earnings[[col]] <- as.numeric(gsub("[\\$,]", "", nyc_weekly_earnings[[col]]))
}

nyc_weekly_earnings <- subset(nyc_weekly_earnings, Year >= 2013)

nyc_weekly_earnings <- nyc_weekly_earnings |> dplyr::select(-Annual)

national_weekly_earnings <- read.csv("Average weekly earnings of all employees, total private, seasonally adjusted-national.csv")

```

```{r}

nyc_weekly_earnings <- pivot_longer(nyc_weekly_earnings, cols = -Year, names_to = "Month", values_to = "Value")
national_weekly_earnings <- pivot_longer(national_weekly_earnings, cols = -Year, names_to = "Month", values_to = "Value")

nyc_weekly_earnings$Location <- "New York City"

national_weekly_earnings$Location <- "National"

weekly_earnings <- rbind(nyc_weekly_earnings, national_weekly_earnings)

weekly_earnings$Date <- as.Date(paste(weekly_earnings$Year, weekly_earnings$Month, "01", sep = "-"), format = "%Y-%b-%d")

weekly_earnings <- weekly_earnings |>
  mutate(PercentChange = c(((Value - lag(Value)) / lag(Value)) * 100))

weekly_earnings <- na.omit(weekly_earnings)

ggplot(weekly_earnings, aes(x = Date, y = Value, color = Location)) +
  geom_line() +
  labs(title = "NYC vs. National Average Weekly Income Over Time",
       x = "Date",
       y = "Average Weekly Income ($)")+
  theme_linedraw()+
  scale_x_date(date_breaks = "2 year", date_labels = "%b %Y")+
  scale_y_continuous(breaks = seq(800, 1450, by = 50))+
  theme(plot.title = element_text(hjust = 0.5))

```

```{r}
weekly_earnings <- weekly_earnings |>
  mutate(PercentChange = c(((Value - lag(Value)) / lag(Value)) * 100))

subsampled_df <- weekly_earnings |> slice(seq(1, n(), by = 2))

subsampled_df <- na.omit(subsampled_df)


ggplot(subsampled_df, aes(x = Date, y = PercentChange, color = Location)) +
  geom_line() +
  labs(title = "Percent Change in NYC vs. National Average Weekly Income Over Time",
       x = "Date",
       y = "Percent Change in Average Weekly Income (%)")+
  theme_linedraw()+
  scale_x_date(date_breaks = "2 year", date_labels = "%b %Y")+
  theme(plot.title = element_text(hjust = 0.5))
```

```{r}

nyc_weekly_earnings$Difference <- nyc_weekly_earnings$Value - national_weekly_earnings$Value

nyc_weekly_earnings$Date <- as.Date(paste(nyc_weekly_earnings$Year, nyc_weekly_earnings$Month, "01", sep = "-"), format = "%Y-%b-%d")

nyc_weekly_earnings <- na.omit(nyc_weekly_earnings)

ggplot(nyc_weekly_earnings, aes(x = Date, y = Difference)) +
  geom_line() +
  labs(title = "Difference Between NYC and National Average Weekly Earnings Over Time",
       x = "Date",
       y = "Average Weekly Income ($)")+
  theme_linedraw()+
  scale_x_date(date_breaks = "2 year", date_labels = "%b %Y")+
  theme(plot.title = element_text(hjust = 0.5))

```

*Weekly hours next (might not include hourly earnings because could be redundant). Then I'll do a parallel coordinate plot for the different sectors and their employment rates. Also don't need to include all three of these graphs, I just thought it was interesting to dive further into depth what was happening. Will also want to do probably just a simple unemployment rate one. *

```{r}

NYC_stats <- read.csv("NYC Employment Statistics.csv")

NYC_stats <- NYC_stats |> dplyr::select(c(YEAR, INDUSTRY_TITLE, JUN))

NYC_stats <- subset(NYC_stats, INDUSTRY_TITLE %in% c("Information", "Telecommunications",
"Financial Activities", "Financial Investments and Related Activities including Financial Vehicles", "Professional and Business Services", "Professional, Scientific, and Technical Services", "Computer Systems Design and Related Services", "Management, Scientific, and Technical Consulting Services", "Scientific Research and Development Services", "Advertising, Public Relations, and Related Services", "Government"))

NYC_stats$NYC_Values <- NYC_stats$JUN

NYC_stats$Year <- NYC_stats$YEAR

NYC_stats <- NYC_stats |> dplyr::select(-JUN)

NYC_stats <- NYC_stats |> dplyr::select(-YEAR)

NYC_stats <- subset(NYC_stats, Year >= 2013)

NYC_stats$NYC_Values <- as.numeric(NYC_stats$NYC_Values)

```

```{r}

telecoms <- read.csv("Telecommunications-national.csv")
SciRe <- read.csv("Scientific research and development-national.csv")
ProSciTech <- read.csv("Professional,scientific,and technical services-national.csv")
ProBuss <- read.csv("Professional and Business Services-national.csv")
ManSciTech <- read.csv("Management, scientific, and technical consulting services-national.csv")
Info <- read.csv("Information(news) - national.csv")
Gov <- read.csv("Government-national.csv")
FinInv <- read.csv("Financial Investments-national.csv")
FinAct <- read.csv("Financial Activities-national.csv")
CompSys <- read.csv("Computer systems design and related services-national.csv")
Advert <- read.csv("Advertising, public relations, and related services-national.csv")


telecoms$INDUSTRY_TITLE <- "Telecommunications"
SciRe$INDUSTRY_TITLE <- "Scientific Research and Development Services"
ProSciTech$INDUSTRY_TITLE <- "Professional, Scientific, and Technical Services"
ProBuss$INDUSTRY_TITLE <- "Professional and Business Services"
ManSciTech$INDUSTRY_TITLE <- "Management, Scientific, and Technical Consulting Services"
Info$INDUSTRY_TITLE <- "Information"
Gov$INDUSTRY_TITLE <- "Government"
FinInv$INDUSTRY_TITLE <- "Financial Investments and Related Activities including Financial Vehicles"
FinAct$INDUSTRY_TITLE <- "Financial Activities"
CompSys$INDUSTRY_TITLE <- "Computer Systems Design and Related Services"
Advert$INDUSTRY_TITLE <- "Advertising, Public Relations, and Related Services"


national_stats <- rbind(telecoms, SciRe, ProSciTech, ProBuss, ManSciTech, Info, Gov, FinInv,
                        FinAct, CompSys, Advert)

national_stats <- national_stats |> dplyr::select(c(Year, INDUSTRY_TITLE, Jun))

national_stats <- pivot_longer(national_stats, cols = -c(Year, INDUSTRY_TITLE), names_to = "Month", values_to = "National_Values")

national_stats <- national_stats |> dplyr::select(-Month)

```

```{r}

NYC_stats_wide <- spread(NYC_stats, key = Year, value = NYC_Values)

national_stats_wide <- spread(national_stats, key = Year, value = National_Values)

NYC_stats_wide$Location <- "New York City"

national_stats_wide$Location <- "National"

national_stats_wide$"2018" <- (national_stats_wide$"2018" / 162210) * 100
NYC_stats_wide$"2018" <- (NYC_stats_wide$"2018" / 4262.878) * 100
national_stats_wide$"2019" <- (national_stats_wide$"2019" / 163073) * 100
NYC_stats_wide$"2019" <- (NYC_stats_wide$"2019" / 4261.786) * 100
national_stats_wide$"2020" <- (national_stats_wide$"2020" / 159900) * 100
NYC_stats_wide$"2020" <- (NYC_stats_wide$"2020" / 4061.283) * 100
national_stats_wide$"2021" <- (national_stats_wide$"2021" / 161192) * 100
NYC_stats_wide$"2021" <- (NYC_stats_wide$"2021" / 4116.222) * 100
national_stats_wide$"2022" <- (national_stats_wide$"2022" / 164002) * 100
NYC_stats_wide$"2022" <- (NYC_stats_wide$"2022" / 4099.68) * 100
national_stats_wide$"2023" <- (national_stats_wide$"2023" / 166951) * 100
NYC_stats_wide$"2023" <- (NYC_stats_wide$"2023" / 4184.884) * 100

stats <- rbind(NYC_stats_wide, national_stats_wide)

ggparcoord(NYC_stats_wide, columns = c(7:12), groupColumn = 1, title = "Potential Industries Labor Force Share in NYC")
```

```{r}

ggparcoord(stats, columns = c(1, 7:12), groupColumn = 13, title = "Comparing NYC Potential Industries Labor Force Percent Share vs. National")
  
```
```{r}

target_occupations = c("Computer and Mathematical Occupations",                                                                  "Computer Occupations",                                                                               
"Computer and Information Analysts",                                                                       
"Computer Systems Analysts",                                                                                
"Information Security Analysts",                                                                            
"Computer and Information Research Scientists",                                                             
"Computer Support Specialists",                                                                         
"Computer Network Support Specialists",                                                            
"Computer User Support Specialists",
"Database Administrators and Architects",
"Database and Network Administrators and Architects",                                                       
"Computer Network Architects",                                                                              
"Database Administrators",                                                                                  
"Database Architects",                                                                                      
"Network and Computer Systems Administrators",                                                              
 "Software and Web Developers, Programmers, and Testers",                                                    
"Computer Programmers",                                                                                     
"Software Developers",                                                                                      
"Software Quality Assurance Analysts and Testers",                                                          
"Web Developers",                                                                                           
"Web and Digital Interface Designers",                                                                      
"Miscellaneous Computer Occupations",                                                                       
"Computer Occupations, All Other",
"Data Scientists",
"Statisticians")
```


```{r}
# Reading the csv files
all_data_2022 = read.csv("all_data_2022.csv")
all_data_2021 = read.csv("all_data_2021.csv")
all_data_2020 = read.csv("all_data_2020.csv")
all_data_2019 = read.csv("all_data_2019.csv")
all_data_2018 = read.csv("all_data_2018.csv")

all_data_2022$YEAR <- "2022"
all_data_2021$YEAR <- "2021"
all_data_2020$YEAR <- "2020"

names(all_data_2019) <- toupper(names(all_data_2019))
all_data_2019$YEAR <- "2019"

names(all_data_2018) <- toupper(names(all_data_2018))
all_data_2018$YEAR <- "2018"
```

```{r}
# Getting a list of NY areas
NY_areas = all_data_2021$AREA_TITLE[all_data_2021$PRIM_STATE == "NY"] |> unique()

# Compiling datasets for NY per year
all_data_2022_NY = all_data_2022 |> subset(PRIM_STATE =="NY")
all_data_2021_NY = all_data_2021 |> subset(PRIM_STATE =="NY")
all_data_2020_NY = all_data_2020 |> subset(PRIM_STATE =="NY")
all_data_2019_NY = all_data_2019 |> subset(AREA_TITLE %in% NY_areas)
all_data_2018_NY = all_data_2018 |> subset(AREA_TITLE %in% NY_areas)

# Adding the missing PRIM_STATE variable
all_data_2019_NY$PRIM_STATE = 'NY'
all_data_2018_NY$PRIM_STATE = 'NY'

# Getting a list of common column names so we can combine the datasets
common_columns = Reduce(intersect, list(names(all_data_2022_NY),names(all_data_2021_NY),names(all_data_2020_NY),names(all_data_2019_NY),names(all_data_2018_NY)))

all_data_NY = rbind(all_data_2022_NY[common_columns], all_data_2021_NY[common_columns], all_data_2020_NY[common_columns], all_data_2019_NY[common_columns], all_data_2018_NY[common_columns])

# Pre-processing the all_data_NY dataset
all_data_NY$A_MEDIAN = gsub(",", "", all_data_NY$A_MEDIAN) |> as.numeric()
all_data_NY$A_MEAN = gsub(",", "", all_data_NY$A_MEAN) |> as.numeric()
all_data_NY$TOT_EMP = gsub(",", "", all_data_NY$TOT_EMP) |> as.numeric()

```

```{r}
# Unifying labels
all_data_NY$OCC_TITLE[grepl("Software Developers", all_data_NY$OCC_TITLE)] = "Software Developers"
all_data_NY$OCC_TITLE[grepl("Data Scientists", all_data_NY$OCC_TITLE)] = "Data Scientists"
all_data_NY$OCC_TITLE[grepl("Database", all_data_NY$OCC_TITLE)] = "Database Administrators and Architects"
```

```{r}
mean_salaries <- all_data_NY[all_data_NY$YEAR == "2022" & all_data_NY$OCC_TITLE %in% target_occupations,] |>
  group_by(OCC_TITLE) |> summarise(Mean_Annual_Salary = mean(A_MEAN))

ggplot(mean_salaries, aes(x = Mean_Annual_Salary, y = reorder(OCC_TITLE, Mean_Annual_Salary ))) +
  geom_point(size = 3) +
  geom_point(data = subset(mean_salaries, OCC_TITLE == "Data Scientists"), color = "red", size = 3) +
  labs(
    title = "Mean Annual Salaries for Tech Positions in NY for 2022",
    x = "Mean Annual Salary in $",
    y = "Occupation"
  )
```

```{r}
focus_occupations = c("Data Scientists" , "Computer Programmers" , "Statisticians" , "Software Developers", "Database Administrators and Architects", "Computer and Information Research Scientists", "Software Quality Assurance Analysts and Testers" )

all_data_NY[all_data_NY$OCC_TITLE %in% focus_occupations,] |>
  group_by(YEAR, OCC_TITLE) |> summarise(Mean_Annual_Salary = mean(A_MEAN , na.rm = TRUE)) |>
  pivot_wider(names_from = YEAR, values_from = Mean_Annual_Salary) |>
  
  ggparcoord(
    columns = 3:6, groupColumn = 1,  scale = "globalminmax", missing = "mean",
    ) +   
    labs(title = "The Mean Annual Salary for Tech Jobs across NY State, 2019-2022 ", 
       x = "Years", 
       y = "Mean Annual Income in $", 
       color = "Occupation")
```

```{r}
all_data_NY[all_data_NY$OCC_TITLE %in% c("Data Scientists", "Software Developers"),] |>
  group_by(YEAR, AREA_TITLE, OCC_TITLE) |> summarise(Mean_Annual_Salary = mean(A_MEAN)) |>
  pivot_wider(names_from = YEAR, values_from = Mean_Annual_Salary) |>
  
  ggparcoord(
    columns = 4:7, groupColumn = 2, showPoints = TRUE,  scale = "globalminmax"
    ) + 
       labs(title = "The Mean Annual Salary for DATA Scientists and Software Developers in Different Regions of NY, 2019-2022 ", 
       x = "Years", 
       y = "Mean Annual Income in $", 
       color = "Occupation")

```

```{r}
all_data_NY[all_data_NY$OCC_TITLE %in% c("Data Scientists", "Software Developers"),] |>
  group_by(YEAR, AREA_TITLE, OCC_TITLE) |> summarise(Mean_TOT_EMP = mean(TOT_EMP)) |>

  ggplot( aes(x=as.numeric(YEAR), y=Mean_TOT_EMP, fill=OCC_TITLE)) + 
    geom_area() +
    facet_wrap(~ AREA_TITLE, scales = "free_y") + 
    theme(plot.margin = margin(0, 0, 0, 0, "cm")) +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
    labs(title = "Total number of Employees per Tech job across NY State Regions", 
       x = "Years", 
       y = "Number of Employees", 
       fill = "Occupation")
```
